Draft: Notes on the revised Hidden weight function

### 2024-10-02 RevisedHiddenWeight.tex
ok

\subsection{Boolean functions and weightwise considerations}

In this part we recall general concepts on Boolean functions and their weightwise properties we use in this article. 
For a deeper introduction on Boolean functions and their cryptographic parameters we refer to the survey of \eg~\cite{Carlet20} and to~\cite{TOSC:CarMeaRot17} for the weightwise properties, also called properties on the slices.
For $k \in [0,n]$ we denote $\Ekn{k}{n}$ the set $\{x\in \F_2^n \, | \, \w(x)=k  \}$ and call it slice of the Boolean hypercube (of dimension $n$). 
Accordingly, the Boolean hypercube is partitioned into $n+1$ slices where the elements have the same Hamming weight.

@@

To denote when a property or a definition is restricted to a slice we use the subscript $k$. 
For example, for a $n$-variable Boolean function $f$ we denote its support $\supp(f)=\{x\in \F_2^n \, | \, f(x)=1  \}$ and we denote $\suppk{k}(f)$ its support restricted to a slice, that is $\supp(f)\cap \Ekn{k}{n}$.

@@

	For $k\in [0,n]$ the function is said balanced on the slice $k$ if $||\suppk{k}(f)|-|\suppk{k}(f+1)| |\le 1$. In particular when $|\Ekn{k}{n}|$ is even $|\suppk{k}(f)|=|\suppk{k}(f+1)|=|\Ekn{k}{n}|/2$.
\end{definition}

@@

\subsection{Symmetric Functions}

The $n$-variable \emph{Boolean symmetric functions} are those that are constant on each slice $\Ekn{k}{n}$ for $k\in [0,n]$. 
This class has been assiduously studied in the context of cryptography, see \eg \cite{IEEE:Carlet04,IEEE:CanVid05,INDO:BraPre05,DM:SarMai07,IEEE:QFLW09,Latin:Meaux19,CCDS:Meaux21,IEEE:CarMea21}.
In this paper we mainly consider two families of symmetric functions, which are both bases of the symmetric functions:

\begin{definition}[Elementary symmetric functions]
Let $i\in [0,n]$, the elementary symmetric function of degree $i$ in $n$ variables, denoted $\sigma_{i,n}$, is the function which ANF contains all monomials of degree $i$ and no monomials of other degrees. 
\end{definition}
\begin{definition}[Slice indicator functions]
 Let $k\in [0,n]$, the indicator function of the slice of weight $k$ is defined as:
 \[\forall  x\in \mathbb{F}_2^{n}, \quad \phikn{k}{n}(x) = 1 \text{ if and only if } \w(x) = k.\]
\end{definition}
We recall that bent functions are those Boolean functions achieving maximal nonlinearity.

@@

\pmnote{Introduce formally HWBF, and weightwise degree d functions}

>>>>

For $v$ a binary vector of length $n$ we denote by $\w(v)$ its Hamming weight: $\w(v)=|\{i\in [1,n] \, | \, x_i=1\}|$.

\pmnote{Preliminaries to adapt}

\subsection{Boolean functions and cryptographic parameters}

In this part we recall general concepts on Boolean functions and their cryptographic properties we use in this article. 
For a deeper introduction on Boolean functions and their cryptographic parameters we refer to the book~\cite{Carlet20} and to~\cite{TOSC:CarMeaRot17} for the weightwise properties, also called properties on the slices.
For $k \in [0,n]$ we denote $\Ekn{k}{n}$ the set $\{x\in \F_2^n \, | \, \w(x)=k  \}$ and call it slice of the Boolean hypercube (of dimension $n$).

%Accordingly, the Boolean hypercube is partitioned into $n+1$ slices where the elements have the same Hamming weight.

@@

@@

%	For $k\in [0,n]$ the function is said balanced on the slice $k$ if $||\suppk{k}(f)|-|\suppk{k}(f+1)| |\le 1$. In particular when $|\Ekn{k}{n}|$ is even $|\suppk{k}(f)|=|\suppk{k}(f+1)|=|\Ekn{k}{n}|/2$.
\end{definition}

@@

\subsection{Symmetric Functions, HWBF and weightwise degree-d functions}

The $n$-variable \emph{Boolean symmetric functions} are those that are constant on each slice $\Ekn{k}{n}$ for $k\in [0,n]$. 
This class has been assiduously studied in the context of cryptography, see \eg \cite{IEEE:Carlet04,IEEE:CanVid05,INDO:BraPre05,DM:SarMai07,IEEE:QFLW09,Latin:Meaux19,CCDS:Meaux21,IEEE:CarMea21}.
In this paper we mainly consider the slice indicator functions.

%\begin{definition}[Elementary symmetric functions]
%Let $i\in [0,n]$, the elementary symmetric function of degree $i$ in $n$ variables, denoted $\sigma_{i,n}$, is the function which ANF contains all monomials of degree $i$ and no monomials of other degrees. 
%\end{definition}
\begin{definition}[Slice indicator functions]\label{def:slice}
 Let $k\in [0,n]$, the indicator function of the slice of weight $k$ is defined as:
 \[\forall  x\in \mathbb{F}_2^{n}, \quad \phikn{k}{n}(x) = 1 \text{ if and only if } \w(x) = k.\]
\end{definition}

%We recall that bent functions are those Boolean functions achieving maximal nonlinearity.

\pmnote{adapt usefull parts:

In~\cite{DAM:GinMea22} the authors study the relationship between weightwise perfectly balanced functions (functions balanced on all the slices, see for example~\cite{TOSC:CarMeaRot17,DCC:LiuMes19,CC:TangLiu19,BFA:MesSuLi21,ACI:ManKuLar22,SSRN:ZLCQZ23,ASIACCS:YCL+23,LATIN:GinMea23}) and weightwise affine functions \ie weightwise degree-$1$ functions. 
In fact weightwise degree-$1$ and degree-$0$ functions have been studied for their cryptographic properties in many works, without the same formalism. 
The weightwise constant functions ($\WD{0}{n}$) are the symmetric functions that have been extensively studied (\eg \cite{IEEE:Carlet04,IEEE:CanVid05,INDO:BraPre05,DM:SarMai07,IEEE:CheLu11,A2C:CarMea19,Latin:Meaux19,CCDS:Meaux21,IEEE:CarMea21}). 
Thereafter, the hidden weight bit function introduced in~\cite{IEEE:Bryant91} is a weightwise degree-$1$ function, the one obtained by fixing $f_0=0$ and $f_k=x_k$ for $k \in [n]$. 
The cryptographic properties of this function have been studied in~\cite{DAM:WCST14}, showing good algebraic properties for this function. 
In~\cite{TOSC:CarMeaRot17}, the bent functions in Propositions $1$ and $2$ are weightwise affine.}

\pmnote{Introduce formally HWBF, and weightwise degree d functions}
\begin{definition}[Hidden Weight Bit Function (HWBF)]\label{def:hwbf}
	Let $n\in \N^*$we call Hidden Weight Bit Function (HWBF) the $n$-variable Boolean function defined by:
	\[h(x)= \left\{ 
	\begin{array}{l l}
	0 & \quad \text{ if } \w(x) =0, \\
	x_k & \quad \text{ if } \w(x)=k.
	\end{array} \right. \]

\end{definition}

\begin{definition}[Weightwise degree-$d$ functions]\label{def:WdegD}
	Let $n\in \N^*$ and for $k\in[0,n]$ $\phikn{k}{n}$ denotes the indicator function of $\Ekn{k}{n}$. 
	An $n$-variable Boolean function $f$, written as $f=\sum_{k=0}^n f_k \phikn{k}{n}$, is called weightwise degree-$d$ if and only if for each $k\in [0,n]$ $f_k$ coincide with a function of degree at most $d$ over $\Ekn{k}{n}$. 
	%	Equivalently, $f$ is weightwise affine if and only if for each $k\in [0,n]$ $\NLk{k}(f)=0$.

	The set of weightwise degree-$d$ functions is denoted by $\WD{d}{n}$.

	%	$f=\sum_{k=0}^n f_k \phikn{k}{n}$ is called weightwise affine if and only if for each $k\in [0,n]$ $f_k$ coincide with an affine function over $\Ekn{k}{n}$.

\end{definition}

\pmnote{add: properties of maj function, HWPF, and some WWL and WWQ}

\pmnote{add the nonlinearity of Dickson's function}

@@
